### 2017-01-23 Main.tex
First version

\documentclass[sigplan, anonymous, review]{acmart}
%\documentclass[prodmode,acmtoplas]{acmsmall}

\usepackage{color}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{stmaryrd}
\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{mathpartir}
\usepackage{wasysym}
\usepackage{url}
\usepackage{algorithm, algpseudocode}
\usepackage{fancyvrb}
%\usepackage[[style=1]{mdframed}
\usepackage{comment}
\usepackage{soul}

\definecolor{light}{gray}{.75}

% \mdfdefinestyle{MyFrame}{%
%     linecolor=white,
% %%    outerlinewidth=2pt,
% %%    roundcorner=20pt,
% %%    innertopmargin=\baselineskip,
% %%    innerbottommargin=\baselineskip,
% %%    innerrightmargin=20pt,
% %%    innerleftmargin=20pt,
%     backgroundcolor=light}

\newcommand{\gbox}[1]{\colorbox{light}{$\!\!#1\!\!$}}

%%\bibliographystyle{plain}

%% for comments
\input{macro-comments}
\input{macros}
\input{Preamble}
\input{lemmas}

% \newenvironment{example}{\par\noindent\textit{Example}\quad}{}

% Metadata Information
%\acmVolume{1}
%\acmNumber{1}
%\acmArticle{1}
%\acmYear{2012}
%\acmMonth{1}

% Document starts
\begin{document}

% Page heads
\markboth{B. Oliveira et al.}{The Implicit Calculus: A New Foundation for Generic Programming}

% Title portion
\title{The Implicit Calculus: A New Foundation for Generic Programming}
\author{Ommitted for anonymous submission}

\begin{abstract}
\input{src/Abstract}
\end{abstract}

\begin{comment}
\category{D.3.2}{Programming Languages}   
                {Language Classifications}
                [Functional Languages, Object-Oriented Languages]
\category{F.3.3}{Logics and Meanings of Programs}   
                {Studies of Program Constructs}
                []

\terms{Languages}

\keywords{
Implicit parameters, type classes, C++ concepts, generic programming,
Haskell, Scala}

\acmformat{Oliveira, B. C. d. S., Schrijvers, T., Choi, W., Lee, W., Yi, K., Wadler, P.
201?. The implicit calculus: a new foundation for generic programming.}

\begin{bottomstuff}
Author's addresses: 
B. C. d. S. Oliveira, Department of Computer Science, Hong Kong University; 
T. Schrijvers, Department of Applied Mathematics and Computer Science, Ghent University;
W. Choi {and} W. Lee {and} K. Yi, \ldots, Seoul National University;
P. Wadler, \ldots, University of Edinburgh.
\end{bottomstuff}
\end{comment}

\maketitle

\input{src/Introduction}

\input{src/Overview}

\input{src/Types}

\input{src/Translation}

\input{src/SourceLang}

%\input{src/Related}

\input{src/Conclusion}

% Appendix
\appendix
\section*{APPENDIX}
\setcounter{section}{1}

%\appendixhead{OLIVEIRA}

% Acknowledgments
\begin{acks}
We are grateful to Ben Delaware, Derek Dreyer, Jeremy Gibbons, Scott
Kilpatrick, eta Ziliani, the members of ROPAS and the
anonymous reviewers for their comments and suggestions.  This work was
partially supported by Korea Ministry of Education, Science and
Technology/Korea Science and Enginering Foundation's ERC grant
R11-2008-007-01002-0, Brain Korea 21, Mid-career Research Program 2010-0022061, and
by Singapore Ministry of Education research grant MOE2010-T2-2-073.
\end{acks}

% Bibliography
\bibliographystyle{ACM-Reference-Format}
\bibliography{sigproc}

%\input{src/Proofs}

% History dates
%\received{February 2007}{March 2009}{June 2009}

\end{document}