### 2017-01-23 src/Abs.tex

Implicit programming that infers values by using type
information has proven its benefits in practice. However, 
years of experience have exposed several limitations in its current practice.
For example, Haskell's type classes, which are the oldest and most
prominent form of implicit programming have several
limitations of its original design: no
ability to control the \emph{scoping} of instances; the \emph{second
  class} nature of types classes compared to types; and type class
constraints being limited to \emph{first-order predicates}.

This paper presents a \textit{implicit calculus}: a powerful and expressive,
yet minimalistic, calculus that provides an implicit programming model
without the limitations of the current practice. The implicit calculus
offers control over scoping, does not suffer from the second class
nature of type classes, and incoorporates two features which
have been discussed in the literature but for which there is
currently no good solution: \emph{higher-order predicates} and
\emph{coherent support for overlapping instances}.  We present the
syntax, operational semantics, type system, type-directed translation
into System F, as well as correctness results. Finally, we show how
Haskell and Scala's implicit programming constructs are translated
into our calculus.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../Main"
%%% End: