### 2017-01-23 src/Termination.tex

\section{Termination of Resolution}

If we are not careful about which rules are made implicit, the recursive
resolution process may not terminate. This section describes how to impose 
a set of modular syntactic restrictions that prevents non-termination. 
%% We have to avoid non-termination of the type checker to ensure soundness.

As an example of non-termination consider 
\begin{equation*}
\myset{
  \myset{\tychar} \To \tyint,
  \myset{\tyint} \To \tychar} \vturns \tyint
\end{equation*}
which loops, using alternatively the first and second rule in the implicit
environment.

The problem of non-termination has been widely studied in the context of
Haskell's type classes, and a set of modular syntactic restrictions
has been imposed on type class instances to avoid non-termination~\cite{fdchr}. 
Adapting these restrictions to our setting, we obtain the following termination
condition.

\defterm

% Haskell's condition is quite severe because the one global scope for all type
% class instances is \textit{open}: more instances can be added later (in other
% modules).  The modularity of the condition already anticipates such future
% additions.
% 
% In contrast, our local scopes are \textit{closed}. Later extensions of the
% program (e.g., new modules) do not affect the existing scopes. Hence, in
% $\ourlang$, termination of resolution coincides with the traditional program
% termination problem. So, alternatively, $\ourlang$  may enforce termination in
% a less stringent manner using available termination checkers like~\cite{approve}.