### 2017-01-24 src/Translation.tex
updating Sections 3 - 4: work in progress

\section{Type-Directed Translation to System F}
\label{sec:trans}

In this section we explain the dynamic semantics of $\ourlang$ in terms
of System F's dynamic semantics, by means of a type-directed translation. 
This translation turns implicit contexts into explicit parameters and
statically resolves all queries, much like Wadler and Blott's dictionary
passing translation for type classes~\cite{adhoc}. 
The advantage of this approach is that we simultaneously provide a meaning to
well-typed $\ourlang$ programs and an effective implementation that resolves
all queries statically.

The translation follows the type system presented in Section~\ref{sec:ourlang}.
The additional machinery that is necessary (on top of the type system)
corresponds to the grayed parts of Figures~\ref{fig:type}, \ref{fig:resolution1} and \ref{fig:resolution2}.

%-------------------------------------------------------------------------------
\subsection{Type-Directed Translation}
Figure~\ref{fig:type} presents the translation rules that convert $\ourlang$
expressions into ones of System F. 
The gray parts of the figure extend the type system with the necessary
information for the translation.

The syntax of System F is as follows: 
{\small
  \[ \begin{array}{llrl}
    \text{Types} & T & ::= & \alpha \mid T \arrow T 
    \mid \forall \alpha. T \\ 
    \text{Expressions} & E & ::=  & x \mid \lambda (x:T) . E \mid E\;E
    \mid \Lambda \alpha . E \mid E\;T 
  \end{array} \]}

The gray extension to the syntax of type environments annotates very
implicit rule type with explicit System F evidence in the form of a 
term variable $x$.

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
\paragraph{Translation of Types}

The function $|\cdot|$ takes 
$\ourlang$ types $\rulet$ to System F types T: 
\begin{eqnarray*}
|\alpha| & = & \alpha \\
|\rulet_1 \arrow \rulet_2| & = & |\rulet_1| \arrow |\rulet_2| \\
|\forall \alpha. \rulet| & = & \forall \alpha. |\rulet| \\
|\rulet_1 \iarrow \rulet_2| & = & |\rulet_1| \arrow |\rulet_2| %\\
\end{eqnarray*}
Its reveals that implicit $\ourlang$ arrows are translated to explicit System F function arrows.

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
\paragraph{Translation of Terms}

The type-directed translation judgment, which extends the typing judgment, is 
\begin{center}
  $\tenv \turns \relation{e}{\rulet}~\gbox{\leadsto E}$
\end{center}
This judgment states that the translation of $\ourlang$ expression $e$ with
type $\rulet$ is System~F expression $E$, with respect to type environment
$\tenv$.

Variables, lambda abstractions and applications
are translated straightforwardly. Perhaps the only noteworthy 
rule is \TyIAbs. This rule associates the type $\rulet_1$ with 
the fresh variable $x$ in the type environment. 
This creates the necessary evidence that can be used by resolutions 
in the body of the rule abstraction to construct System F terms of type $|\rulet_1|$.

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
\paragraph{Resolution}
The more interesting part of the translation happens when resolving queries.
Queries are translated by rule $\TyQuery$ using the auxiliary resolution
judgment $\vturns$:
\begin{equation*}
\tenv \vturns \rulet~\gbox{\leadsto E}
\end{equation*}
which is shown, in deterministic form, in Figure~\ref{fig:resolution2}.
The translation builds a System F term as evidence for the resolution.
The mechanism that builds evidence dualizes the process of peeling off
abstractions and universal 
quantifiers. The rule \mylabel{R-IAbs}~wraps a lambda binder with a fresh variable 
$x$ around a System F expression $E$, which is generated from the resolution 
for the head of the rule ($\rulet_2$). The rule \mylabel{R-TAbs}~wraps a type lambda binder 
around the System F expression resulting from the resolution of $\rulet$.

TODO

For simple types \mylabel{R-Simp}, evidence $E_1$ is retrieved from matching 
$\type$ and then used in the resolution process of the simple type 
$\type$ with rule $\turns_\downarrow$:

\begin{center}
$\bar{\alpha};\env; \rulet~\gbox{\leadsto E_1} \turns_\downarrow \type~\gbox{\leadsto E_2}$
\end{center}

The rules \mylabel{I-IAbs}~and \mylabel{I-TAbs}~are the most interesting in the translation. 
In rule \mylabel{I-IAbs}~we need evidence for $\rulet_2$ to build evidence for $\type$, 
but all we have is evidence $E_1$ for $\rulet_1 \iarrow \rulet_2$. However, we can 
construct evidence for $\rulet_2$, by generating evidence $E_2$ for $\rulet_1$
and then simply applying $E_1$ to $E_2$. Rule \mylabel{I-TAbs}~is similar: we need 
evidence for $\rulet$, but all we have is evidence $E_1$ for $\forall \alpha. \rulet$. 
To create evidence for $\rulet$ we can substitute $\alpha$ by some type 
$\rulet'$ in $\rulet$ and generate a term which is the type application of 
$E_1$ to $|\rulet'|$.

Finally, matching ($\elookup{\env}{\type}$) also needs to be
extended with evidence generation, but this extension is
straightforward.

\paragraph{Meta-Theory} The type-directed translation of $\ourlang$ to System F exhibits a number
of desirable properties.

\begin{theorem}[Type-preserving translation]\label{thm:type:preservation} Let $e$ be an $\ourlang$
  expression, $\rulet$ be a type, $\tenv$ a type environment and $E$ be a System F expression. If
  $\tenv \turns \relation{e}{\rulet} \leadsto E$, then $|\tenv| \turns \relation{E}{|\rulet|}$.
\end{theorem}
Here we define the translation of the type environment form $\ourlang$ to System F as:
\begin{equation*}
\begin{array}{rcl@{\hspace{2cm}}rcl}
|\epsilon| & = & \epsilon & |\tenv,\alpha| & = & |\tenv|, \alpha \\
|\tenv,x : \rulet| & = & |\tenv|, x : |\rulet| &
|\tenv, \rulet \leadsto x| & = & |\tenv|, x : |\rulet|
\end{array}
\end{equation*}

An important lemma in the theorem's proof is the type preservation of 
resolution.
\begin{lemma}[Type-Preserving Resolution]
Let $\tenv$ be a type environment, $\rulet$ be a type and E be a System F expression.
If $\tenv \vturns \rulet \leadsto E$, then $|\tenv| \vdash E : |\rulet|$.
\end{lemma}
% Both the theorem and the lemma are proven in Appendix~\ref{proof:preservation}.

Moreover, we can express three key properties of Figure~\ref{fig:resolution2}'s
definition of resolution in terms of the generated evidence.
\begin{lemma}[Determinacy]
The generated evidence of resolution is uniquely determined.
\[\forall \tenv, \rulet, E_1, E_2: \quad\quad \tenv \ivturns \rulet \leadsto E_1 ~\wedge~ \tenv \ivturns \rulet \leadsto E_2 \quad\Rightarrow\quad E_1 = E_2 \]
\end{lemma}
\begin{lemma}[Soundness]
Figure~\ref{fig:resolution2}'s definition of resolution is sound (but
incomplete) with respect to Figure~\ref{fig:resolution1}'s definition.
\[\forall \tenv, \rulet, E: \quad\quad \tenv \ivturns \rulet \leadsto E \quad\Rightarrow\quad \tenv \vturns \rulet \leadsto E \]
\end{lemma}
\begin{lemma}[Coherence]
Resolution is stable under substitution.
\[\forall \tenv,\alpha,\tenv',\sigma,\rulet, E: \quad\quad 
\tenv,\alpha,\tenv' \ivturns \rulet \leadsto E \quad\wedge\quad \tenv \vdash \sigma
\quad\Rightarrow\quad 
\tenv,\tenv'[\sigma/\alpha] \ivturns \rulet[\sigma/\alpha] \leadsto E[|\sigma|/\alpha] \]
\end{lemma}

%-------------------------------------------------------------------------------
\subsection{Evidence Generation in the Algorithm}

The evidence generation in Figure~\ref{fig:algorithm} is largely similar to
that in the deterministic specification of resolution in
Figure~\ref{fig:resolution2}.

% The main difference, and complication, is due to the fact that the evidence for
% type instantiation and recursive resolution is needed before these operations
% actually take place, as the algorithm has to postpone them. For this reason, the
% algorithm first produces placeholders that are later substituted for the actual
% evidence.
% 
% The central relation is $\rulet; \bar{\rulet}; \bar{\alpha} \gbox{; \bar{\omega};
% E}\turns_{\mathit{match}} \tau \hookrightarrow \bar{\rulet}' \gbox{;
% \bar{\omega}'; E'}$. It captures the matching instantiation of context type
% $\rulet$ against simple type $\tau$.  The input evidence for $\rulet$ is $E$, and
% the output evidence for the instantiation is $E'$. The accumulating parameters
% $\bar{\alpha}$ and $\bar{\rulet}$ denote that the instantiation of type variables
% $\bar{\alpha}$ and the recursive resolution of $\bar{\rulet}$ have been
% postponed. We use the $\bar{\alpha}$ themselves as convenient placeholders for
% the instantiating types, and we use the synthetic $\bar{\omega}$ as placeholders for
% the evidence of the $\bar{\rulet}$. The rules \mylabel{MTC-Abs} and \mylabel{MTC-Abs} 
% introduce these two kinds of placeholders in the evidence. The former kind, $\bar{\alpha}$,
% are substituted in rule \mylabel{MTC-Simp} where the actual type instantiatons $\bar{\theta}$
% are computed. The latter kind, $\bar{\omega}$, are substituted later in rule \mylabel{Alg-Simp} where
% the recursive resolutions take place.

With the evidence we can state the correctness of the algorithm.
\begin{theorem}[Partial Correctness]
Let $\tenv$ be a type environment, $\rulet$ be a type and E be a System F expressions.
Assume that $\epsilon \vdash_{\mathit{unamb}} \rulet$ and also $\forall \rulet_i \in \tenv: \epsilon \vdash_{\mathit{unamb}} \rulet_i$.
Then $\tenv \ivturns \rulet \leadsto E$ if and only if $\tenv \vdash_{\mathit{alg}} \rulet \leadsto E$,
provided that the algorithm terminates.
\end{theorem}

%-------------------------------------------------------------------------------
\subsection{Dynamic Semantics}
Finally, we define the dynamic semantics of $\ourlang$ as the composition of
the type-directed translation and System F's dynamic semantics.  Following
Siek's notation~\cite{systemfg}, this dynamic semantics is:
\[ \mathit{eval}(e) = V \quad\quad \textit{where } \epsilon \turns \relation{e}{\rulet} \leadsto E \textit{ and } E \rightarrow^* V  \]
with $\rightarrow^*$ the reflexive, transitive closure of System F's standard single-step call-by-value reduction relation (see \cite[Chapter 23]{tapl}).

Now we can state the conventional type safety theorem for $\ourlang$:
\begin{theorem}[Type Safety]
If $\epsilon \turns \relation{e}{\rulet}$, then $\mathit{eval}(e) = V$ for
some System F value $V$.
\end{theorem}
The proof follows trivially from Theorem~\ref{thm:type:preservation}.