### 2017-04-19 response.txt
draft response

* Presentation Improvements

  Thank you for the suggestions to improve the presentation.  We will
definitely take them into account. In particular we are very happy with the
comment on making a punchline at the end of Section 2 on stability as the key
idea; we agree that this will be very helpful to better convey the key idea to
readers.

  Tracking of variables, image what happens if we don't treat them spefically
  and consider this example resolution query

  a, a -> a |-_r forall b. b -> b

* Bugs in the typing rules

  We do indeed intentionally abuse the notation (we would not call
  that bugs), but are happy to reformulate as you suggest to clarify
  the intended meaning.

  Should we say something about Show?

* The intended type of bad on p. 5 is actually bad :: a -> a, 
  and not bad :: Trans a => a -> a, because we want to force
  the resolution of Trans at the definition of bad instead of
  the call sites of bad.

* Naming the implicit parameters only works if you want to directly
  refer to them. It does not work in the case of recursive resolution,
  e.g. Int => Bool, Int |- Bool, where multiple rules have to be combined.
  In such cases the system figures out how to combine the rules; the programmer
  does not have to write it himself.

* Provide example for \beta >_\tenv  \alpha

REVIEW 3

* We will give more practical examples of implicits, but this paper is mostly
  about restrictions to obtain coherence. The two cited papers

	B. C. d. S. Oliveira, A. Moors, and M. Odersky. 2010. Type classes as objects
	and implicits. In OOPSLA. ACM, New York, NY, USA.

	B. C. d. S. Oliveira, T. Schrijvers, W. Choi, W. Lee, and K. Yi. 2012.  e
	Implicit Calculus: A New Foundation for Generic Programming. In PLDI ’12. ACM,
	New York, NY, USA.

  provide many examples to motivate implicits.