### 2017-04-20 response.txt

* Presentation Improvements

  Thank you for the suggestions to improve the presentation.  We will
  definitely take them into account. In particular we are very happy with the
  comment on making a punchline at the end of Section 2 on stability as the key
  idea; we agree that this will be very helpful to better convey the key idea to
  readers.

  We would like to clarify the need for tracking the type variables referred
  to on p.13. Consider for example resolving the type

		ρ = b.b → b

  against the environment

  		Γ0 = ∀c.c → c ~> id, a, a → a ~> f

  i.e.

	     	Γ0 ⊢ ρ

  Imagine what would happen if we did not track the type variables. Firstly,
  using rule R-TAbs we would extend the environment with b

  		Γ1 = Γ0, b

  and resolve the simple type

	     	Γ1 ⊢ b → b

  which would get stuck on a → a when considering which rule matches in Γ1. In particular,
  L-RuleMatch does not apply because a → a is different from b → b. Also L-RuleNoMatch does
  not apply because a → a and b → b unify under the substitution [b := a]; thus skipping 
  rule a → a would be deemed unstable.

  However, there is an error in the above reasoning. Code inlining cannot
  result in b being instantiated to a, because b is not free in the environment 
  Γ0, which is the environment at the point in the code where the query happens.
  To avoid this reasoning error, we keep track of the variables that are free
  in the environment in Γ0. These are the variables that can be substituted, while
  the variables added to the environment by rule R-TAbs cannot be substituted.

  We will integrate this example in the text.

* Bugs in the typing rules

  We do indeed intentionally abuse the notation (we would not call
  that bugs), but are happy to reformulate as you suggest to clarify
  the intended meaning.

  TODO: Should we say something about Show?

* The intended type of bad on p. 5 is actually bad :: a -> a, 
  and not bad :: Trans a => a -> a, because we want to force
  the resolution of Trans at the definition of bad instead of
  the call sites of bad.

* Naming the implicit parameters only works if you want to directly
  refer to them. It does not work in the case of recursive resolution,
  e.g. Int => Bool, Int |- Bool, where multiple rules have to be combined.
  In such cases the system figures out how to combine the rules; the programmer
  does not have to write it himself.

* TODO: Provide example for \beta >_\tenv  \alpha

REVIEW 3

* We will give more practical examples of implicits, but this paper is mostly
  about restrictions to obtain coherence. The two cited papers

	B. C. d. S. Oliveira, A. Moors, and M. Odersky. 2010. Type classes as objects
	and implicits. In OOPSLA. ACM, New York, NY, USA.

	B. C. d. S. Oliveira, T. Schrijvers, W. Choi, W. Lee, and K. Yi. 2012.  e
	Implicit Calculus: A New Foundation for Generic Programming. In PLDI ’12. ACM,
	New York, NY, USA.

  provide many examples to motivate implicits.