### 2017-06-06 Main.tex
jfp version: first port

\documentclass[acmlarge, authorname]{acmart}
%\documentclass[prodmode,acmtoplas]{acmsmall}

\usepackage{color}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{stmaryrd}
\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{mathpartir}
\usepackage{wasysym}
\usepackage{url}
\usepackage{algorithm, algpseudocode}
\usepackage{fancyvrb}
%\usepackage[[style=1]{mdframed}
\usepackage{comment}
\usepackage{soul}

\definecolor{light}{gray}{.75}

% \mdfdefinestyle{MyFrame}{%
%     linecolor=white,
% %%    outerlinewidth=2pt,
% %%    roundcorner=20pt,
% %%    innertopmargin=\baselineskip,
% %%    innerbottommargin=\baselineskip,
% %%    innerrightmargin=20pt,
% %%    innerleftmargin=20pt,
%     backgroundcolor=light}

\newcommand{\gbox}[1]{\colorbox{light}{$\!\!#1\!\!$}}
\newcommand{\plw}[1]{[{\color{blue}\text{PLW:}} #1]}

%%\bibliographystyle{plain}

%% for comments
\input{macro-comments}
\input{macros}
\input{Preamble}
\input{lemmas}

% \newenvironment{example}{\par\noindent\textit{Example}\quad}{}

% Metadata Information
%\acmVolume{1}
%\acmNumber{1}
%\acmArticle{1}
%\acmYear{2012}
%\acmMonth{1}

% Document starts
\bibliographystyle{ACM-Reference-Format}
\citestyle{acmauthoryear}
\begin{document}

% Page heads
% Title portion
\title{\name: Deterministic and Coherent Implicits}
\author{Tom Schrijvers}
\affiliation{KU Leuven}
\author{Bruno C.d.S. Oliveira}
\affiliation{The University of Hong Kong}
\author{Philip Wadler}
\affiliation{University of Edinburgh}

\begin{abstract}
\input{src/Abstract}
\end{abstract}
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10003752.10010124.10010125.10010130</concept_id>
<concept_desc>Theory of computation~Type structures</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011006.10011008.10011009.10011012</concept_id>
<concept_desc>Software and its engineering~Functional languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Theory of computation~Type structures}
\ccsdesc[500]{Software and its engineering~Functional languages}

\setcopyright{none}
\maketitle

\input{src/Introduction}

\input{src/Overview}

\input{src/Types}

\input{src/Translation}

%\input{src/SourceLang}

\input{src/Related}

\input{src/Conclusion}

%\appendixhead{OLIVEIRA}

% Acknowledgments
\begin{acks}
We are grateful to Ben Delaware, Derek Dreyer, Jeremy Gibbons, Scott
Kilpatrick, eta Ziliani, the members of ROPAS and the
anonymous reviewers for their comments and suggestions.  This work was
partially supported by Korea Ministry of Education, Science and
Technology/Korea Science and Enginering Foundation's ERC grant
R11-2008-007-01002-0, Brain Korea 21, Mid-career Research Program 2010-0022061, and
by Singapore Ministry of Education research grant MOE2010-T2-2-073.
\end{acks}

\newpage
% Bibliography
\bibliography{papers}

\newpage
% Appendix
\appendix
\section*{APPENDIX}
\setcounter{section}{1}

\input{src/Proofs}

% History dates
%\received{February 2007}{March 2009}{June 2009}

\end{document}